---
title: "Elastic Net Results"
author: "Hillary Heiling"
date: "02/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



```{r, message=FALSE, warning=FALSE}
# Load relevant libraries
library(stringr)
library(xtable)
library(ggplot2)
library(reshape2)
```


```{r}
# Extract relevant information from simulation results
extract = function(path, sims, sim_type,
                   q = 101, p_true = 11, r_true_vec = 3, trace = 0,
                   beta_true = rep(1,length(sims))){
  
  # Results of r estimation procedure
  ## Percent of times too small, too large, or correct
  rest_pseudo = matrix(0, nrow = length(sims), ncol = 3)
  rownames(rest_pseudo) = sim_type
  colnames(rest_pseudo) = c("Too_small_Pct","Correct_Pct","Too_Large_Pct")
  ## Average estimate of r
  rest_avg_pseudo = numeric(length(sims))
  names(rest_avg_pseudo) = sim_type
  
  rest_lst = list()
  
  # Save summary output in out_mat
  ## labs = labels of columns of out_mat
  labs = c(str_c("Beta",1:(p_true-1)),"TP % Fixef","FP % Fixef",
             "TP % Ranef","FP % Ranef",
             "Median Time (hrs)","Abs. Dev. (Mean)")
  out_mat = matrix(0, nrow = length(sims), ncol = length(labs))
  colnames(out_mat) = labs
  rownames(out_mat) = sim_type
  
  # Save additional variance and fixed effect information
  vars_lst = list()
  beta_lst = list()
  
  r_input_lst = list()
  y_lst = list()
  
  # Save eigenvalues and true variances of random effect covariance matrix used in simulations
  eigen_values = matrix(0, nrow = length(sims), ncol = max(r_true_vec))
  rownames(eigen_values) = sim_type
  vars_true = matrix(0, nrow = length(sims), ncol = p_true)
  rownames(vars_true) = sim_type
  
  # Average values for all fixed effects selected in best model as well as
  ## average variances selected in best models
  beta_all_avg = matrix(0, nrow = length(sims), ncol = q)
  rownames(beta_all_avg) = sim_type
  colnames(beta_all_avg) = str_c("Beta",0:(q-1))
  
  vars_all_avg = beta_all_avg
  colnames(vars_all_avg) = str_c("Var",0:(q-1))
  
  # Save bias information
  abs_dev_lst = list()
  # Save time information
  time_lst = list()
  # Save warning information
  warn_lst = list()
  
  for(m in 1:length(sims)){
    
    files = list.files(path = str_c(path, sims[m]), full.names = T)
    if(trace == 1){
      cat("completed replicates: ", length(files), "\n")
    }
    
    if(length(r_true_vec) == 1){
      r_true = r_true_vec
    }else if(length(r_true_vec) > 1){
      r_true = r_true_vec[m]
    }
    
    time_mat = numeric(length(files))
    EM_iter_lst = list()
    r_est = numeric(length(files))
    warn_sim = list()
    
    beta_mat = matrix(0, nrow = length(files), ncol = q)
    abs_dev = numeric(length(files)) # Mean absolute deviation
    vars_mat = matrix(0, nrow = length(files), ncol = q)
    presc_mat = matrix(0, nrow = length(files), ncol = q)
    EM_iter = NULL
    opt_res = NULL
    select_res = list()
    r_input = numeric(length(files))
    y_summary = matrix(NA, nrow = length(files), ncol = 6)
    
    for(f in 1:length(files)){
      # load output list object
      load(files[f])
      beta_mat[f,] = output$coef_mat
      beta_vec = output$coef_mat[,c(2:p_true)] # Ignore intercept in absolute deviation calculation
      beta_non0 = beta_vec[which(beta_vec != 0)]
      if(length(beta_non0) >= 1){
        abs_dev[f] = mean(abs(beta_non0 - rep(beta_true[m],times=length(beta_non0))))
      }else{
        abs_dev[f] = NA
      }
      vars_mat[f,] = output$vars_mat
      presc_mat[f,] = output$presc_mat
      opt_res = rbind(opt_res, output$results_optim[[1]])
      EM_iter = rbind(EM_iter,output$selection_results[,"EM_iter"])
      time_mat[f] = output$time_mat[1,3]
      select_res[[f]] = output$selection_results
      if(!is.null(output$r_est)){
        r_est[f] = output$r_est
      }
      y_tmp = output$y
      if(!is.null(y_tmp)){
        y_summary[f,] = summary(y_tmp)[1:6]
        if(f == 1){
          colnames(y_summary) = names(summary(y_tmp))[1:6]
        }
      }
      
      if(!is.null(output$r_input)){
        r_input[f] = output$r_input
      }else{
        r_input[f] = NA
      }
      
      warn_sim[[f]] = output$warn_list
      # print(output$warn_list)
      
    } # End f for loop
    
    warn_lst[[m]] = warn_sim
    
    abs_dev_lst[[m]] = abs_dev
    
    EM_iter_lst[[m]] = EM_iter
    
    time_lst[[m]] = time_mat
    
    # True and false positives - fixed effects, random effects,
    # "TP PreSc","FP Presc","Med Presc"
    out = numeric(length(labs))
    names(out) = labs
    
    
    for(i in 2:p_true){
      out[i-1] = mean(beta_mat[which(beta_mat[,i] != 0),i])
    }
    for(i in 1:q){
      # Note: will be NA if never != 0
      beta_all_avg[m,i] = mean(beta_mat[which(beta_mat[,i] != 0),i])
      vars_all_avg[m,i] = mean(vars_mat[which(vars_mat[,i] != 0),i])
    }
    out[p_true] = sum(beta_mat[,c(2:p_true)] != 0) / length(files) * (1 / (p_true-1) * 100)
    out[p_true+1] = sum(beta_mat[,-c(1:p_true)] != 0) / length(files) * (1 / (q - p_true) * 100)
    out[p_true+2] = sum(vars_mat[,c(2:p_true)] != 0) / length(files) * (1 / (p_true-1) * 100)
    out[p_true+3] = sum(vars_mat[,-c(1:p_true)] != 0) / length(files)  * (1 / (q - p_true) * 100)
    out[p_true+4] = median(time_mat / 3600)
    out[p_true+5] = mean(abs_dev, na.rm = TRUE)
    
    out_mat[m,] = out
    
    # summary of r estimation using pseudo random effects estimates
    rest_pseudo[m,1] = sum(r_est < r_true) / length(files) * 100
    rest_pseudo[m,2] = sum(r_est == r_true) / length(files) * 100
    rest_pseudo[m,3] = sum(r_est > r_true) / length(files) * 100
    
    rest_avg_pseudo[m] = mean(r_est)
    
    rest_lst[[m]] = r_est
    vars_lst[[m]] = vars_mat
    beta_lst[[m]] = beta_mat
    r_input_lst[[m]] = r_input

    y_lst[[m]] = y_summary
    
    eigen_values[m,] = output$eigen_vals[1:max(r_true_vec)]
    B_true = output$B_true
    vars_true[m,] = diag(B_true %*% t(B_true))
  }
  
  
  
  out_mat = round(out_mat, digits = 2)
  rest_pseudo = round(rest_pseudo, digits = 2)
  rest_avg_pseudo = round(rest_avg_pseudo, 2)
  
  vars_val = round(diag(output$B_true %*% t(output$B_true)), digits = 2)
  
  return(list(out_mat = out_mat, 
              EM_iter = EM_iter_lst, 
              rest_pseudo = rest_pseudo,
              rest_avg_pseudo = rest_avg_pseudo,
              vars_val = vars_val, rest_lst = rest_lst,
              vars_lst = vars_lst, beta_lst = beta_lst,
              beta_all_avg = beta_all_avg, vars_all_avg = vars_all_avg,
              r_input = r_input_lst,
              y_lst = y_lst,
              eigen_values = eigen_values,
              vars_true = vars_true,
              time_lst = time_lst,
              abs_dev_lst = abs_dev_lst,
              warn_lst = warn_lst))
  
}
```

# glmmPen\_FA Elastic Net Simulation Results

General settings:

* 100 total predictors, 10 true predictors with non-zero fixed and random effects
* Fixed effects: $\beta = 1$
* $r=3$, $B$ = moderate (see "Paper_Results.Rmd" for details)
* $r$ estimated using GR method
* 100 simulation replicates per setting

Correlation settings:

* Equal correlation among all covariates, value 0.2
* Equal correlation among all covariates, value 0.4
* Equal correlation among all covariates, value 0.6
* CS: correlation among covariates based on correlation matrix from the case study dataset

Penalization settings:

* minimum value = 0.05 x max value
* $\alpha = \{0.1,0.3,0.5,0.8,1.0\}$

```{r}
# True number of predictors with non-zero fixed and random effects in these simulations
p_true = 11
```

```{r}
# Note: For others interested in re-creating the output here, replace the code below with the line:
# load("Paper_Results/select100_glmmPen_FA_EN03.RData")
# This will load the 'res' list object with the relevant output

# Path to full simulation results
path = "~/Longleaf/Project2/Simulation_Results/select100_glmmPen_FA_EN03/"
# Sub-folders with individual simulation results
sims = str_c("sim",1:20)
# Description of individual simulation results
sim_type = str_c("cor_",rep(c("0.2","0.4","0.6","CS"),times = 5),
                 "_alpha_",rep(c(0.1,0.3,0.5,0.8,1.0), each = 4))
res = extract(path = path, sims = sims, sim_type = sim_type,
              q = 101, p_true = 11, r_true_vec = 3, 
              trace = 0, beta_true = rep(1, length(sims)))
save(res, file = "Paper_Results/select100_glmmPen_FA_EN03.RData")
```


### Comparison of r estimation on psuedo random effects

Average $r$ and percent of times r underestimated, overestimated, and estimated correctly

```{r}
# Re-order so that group by correlation type, then by alpha value
idx = NULL
for(i in 1:4){
  idx = c(idx, seq(from = i, to = length(sims), by = 4))
}
data.frame(Avg_r = res$rest_avg_pseudo, res$rest_pseudo)[idx,]
```

### Algorithm Performance

True and false positive rates:

```{r}
res$out_mat[idx,-c(1:(p_true-1))]
```



### Tables used in paper:

```{r}
# latex tables

# True/False Positives and Timing
tab_tmp = res$out_mat[idx,-c(1:(p_true-1))]
tab_out = data.frame(Cor = rep(c("0.2","0.4","0.6","CS"), each = 5),
                     Pi = rep(c("0.1","0.3","0.5","0.8","1.0"), times = 4),
                     tab_tmp)
# rownames(tab_out) = NULL
print(xtable(tab_out))

# r estimate results
r_out = data.frame(Cor = rep(c("0.2","0.4","0.6","CS"), each = 5),
                   Pi = rep(c("0.1","0.3","0.5","0.8","1.0"), times = 4),
                   Avg_r = res$rest_avg_pseudo[idx], res$rest_pseudo[idx,])
rownames(r_out) = NULL
print(xtable(r_out, digits = c(0,0,0,2,0,0,0))) 
```

### Warnings

During a handful of variable selection procedures, the ...

In Linux window, check:

```
grep small Project2/Rout/EN_03_*.Rout
```

Occurrences of small intercept variance error: 

* Sim 12: Correlation based on case study, alpha = 0.5
    + Replicate 44
* Sim 16: Correlation based on case study, alpha = 0.8
    + Replicate 9, 22, 31, 91
* Sim 20: Correlation based on case study, alpha = 1.0
    + Replicate 16, 25

### Correlation D

Summary of absolute values of correlations used in analysis:

```{r}
load(sprintf("%s/sim4/Output_001.RData",path))
tmp = abs(output$cor_use)
tmp[lower.tri(tmp)] = NA
tmp_vec = c(tmp)
tmp_vec = tmp_vec[which((tmp_vec != 1) & !is.na(tmp_vec))]
summary(tmp_vec)
hist(tmp_vec)
```


The End